---
title: "SCHACE-tutorial"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SCHACE-tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
The **SCHACE** package is created for implementing the Simultaneous CHange-point
detection And Curve Estimation (SCHACE) method, which was proposed for detecting jumps in data and capturing nonlinear
trends between jumps in the mean curve. This tutorial shows how to use all the functions in this package with the simulated data for curve fitting and change-point detection. This package depends on R packages **glmnet**, **splines**, and **cvTools**, so you need to install these packages before running the R code in this tutorial.

We are going to solve this optimization problem:
$$\textrm{min}_{\boldsymbol{\beta} \in \mathbb{R}^{N-1}, \boldsymbol{\gamma} \in \mathbb{R}^{df}} \quad \frac{1}{2}\|\mathbf{y} - \mathbf{X}\boldsymbol{\beta} -  \mathbf{Z}\boldsymbol{\gamma}\|_2^2 + \lambda \|\boldsymbol{\beta}\|_1$$

Firstly, let's generate the simulated data from function **burt** (Abramovich et al, 2007).
```{r setup}
data_burt <- function(n = 256, range = 1, rsnr = 4){
  
  sample_x <- seq(0, range, length.out = n)
  true_y <- 20*sample_x*cos(16*sample_x^(1.2)) - 20*as.numeric(identity(sample_x < 0.5))
  
  sdf <- sd(true_y)
  sigma <- sdf/rsnr
  
  epsilon <- rnorm(n, 0, sigma)
  
  sample_y <- true_y + epsilon
  
  position = which(as.numeric(identity(sample_x < 0.5) == 0) > 0)[1] - 1
  true_CP <- sample_x[position]
  
  data = list("true_y" = true_y, "sample_x" = sample_x, "sample_y" = sample_y, "true CPs" = true_CP, "noise" = sigma)
  return(data)
  
}

set.seed(15)
data_sample <- data_burt(rsnr = 4)
y <- data_sample$"sample_y"
```
The plot of the data sequence:
```{r fig1, fig.height = 5, fig.width = 5, fig.align = "center"}
plot(data_sample$"sample_x", y, xlab = 'x', ylab = 'y', pch = 20, col = "gray")
```

## User-specified parameters
If user specifies the values of $\lambda$ and degrees of freedom (df) of B-splines, say $\lambda = 0.3$ and df = 10, the example code will be
```{r}
library(SCHACE)
results <- main.SCHACE(
  y,
  folds = 3,
  lambda = 0.3,
  Bdf = 10,
  )
predicted_y <- results$"predicted y"
results$"locations of detected CPs"
```
The estimated locations of change points are
```{r}
results$"locations of detected CPs"
```
The graph shows the fitted curves (blue line) and locations of detected change points (orange dashed lines).
```{r fig2, fig.height = 5, fig.width = 5, fig.align = "center"}
plot(data_sample$"sample_x", y, xlab = 'x', ylab = 'y', pch = 20, col = "gray")
lines(data_sample$"sample_x", predicted_y, col = "blue", lty = 1, lwd = 2)
abline(v=results$"locations of detected CPs"/256, col="orange", lty=2, lwd=2)
```

## Select tuning parameters by Cross-validation procedure

If we want to tune parameters by 3-fold cross-validation procedure with trimmed L2-error, where 10% of the largest squared differences of the observed and predicted values of $y$ are trimmed, the example code could be
```{r}
results_CV <- main.SCHACE(
  y,
  folds = 3,
  Bdf_set = seq(3, 15, 1),
  tuning = "crossvalidation",
  methods = "Trim",
  clambda = "lambdamin",
  percent = 0.1
  )
```
```{r fig3, fig.height = 5, fig.width = 5, fig.align = "center"}
plot(data_sample$"sample_x", y, xlab = 'x', ylab = 'y', pch = 20, col = "gray")
lines(data_sample$"sample_x", results_CV$"predicted y", col = "blue", lty = 1, lwd = 2)
abline(v=results_CV$"locations of detected CPs"/256, col="orange", lty=2, lwd=2)
```

## Select tuning parameters by Information Criteria

If we want to tune parameters by extended Bayesian Information Criteria (eBIC), the example code could be
```{r}
results_BIC <- main.SCHACE(
  y,
  tuning = "IC",
  IC = "BIC.Chen"
  )
```
```{r fig4, fig.height = 5, fig.width = 5, fig.align = "center"}
plot(data_sample$"sample_x", y, xlab = 'x', ylab = 'y', pch = 20, col = "gray")
lines(data_sample$"sample_x", results_BIC$"predicted y", col = "blue", lty = 1, lwd = 2)
abline(v=results_BIC$"locations of detected CPs"/256, col="orange", lty=2, lwd=2)
```
